### 2018-01-16 ManuscriptGeophysics2018.tex
Include feeback KK

% \documentclass[manuscript]{geophysics}
\documentclass[paper,twocolumn,twoside]{geophysics}
% \documentclass[paper]{geophysics}
% \documentclass[paper, revised]{geophysics}
% \documentclass[manuscript,revised]{geophysics}

%%fakesection ===    PACKAGES & DEFINITIONS    ===
% \usepackage{showframe}

%% ~ ADDITIONAL PACKAGES TO GEOPHYSICS.CLS
\usepackage[USenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{upquote}
\usepackage[strings]{underscore}
\usepackage[pdftex, final]{hyperref}
\hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}

% Figures
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\renewcommand{\figdir}{./figures}

% Own commands
\usepackage{xspace}
\newcommand{\empymod}{\texttt{empymod}\xspace}
\newcommand{\mr}[1]{\mathrm{#1}}

\begin{document}

\title{A tool for designing digital filters for the Hankel and Fourier transforms in
potential, diffusive, and wavefield modeling}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\ms{GEO-2018-????}

\address{
\footnotemark[1]Instituto Mexicano del Petróleo,
Eje Central Lázaro Cárdenas Norte 152,
Col. San Bartolo Atepehuacan C.P. 07730,
Ciudad de México, México,
E-mail: \href{mailto:dieter@werthmuller.org}{Dieter@Werthmuller.org};
\footnotemark[2]Lamont-Doherty Earth Observatory, Columbia University,
305C Oceanography,
61 Route 9W, PO Box 1000,
Palisades NY 10964-8000 US,
E-mail: \href{mailto:KKey@ldeo.columbia.edu}{KKey@ldeo.columbia.edu};
\footnotemark[3]TU Delft,
Building 23,
Stevinweg 1 / PO-box 5048,
2628 CN Delft,
E-mail: \href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl}.}

\author{%
Dieter Werthmüller\footnotemark[1], %
Kerry Key\footnotemark[2], and %
Evert C.\ Slob\footnotemark[3]%
}

\footer{}
\lefthead{Werthmüller et al.}
\righthead{Digital filter designing tool}

\maketitle

%%fakesection ===    ABSTRACT    ===
\begin{abstract} % 1-2 sentence(s) each
%
% (1) PRINCIPAL OBJECTIVES AND SCOPE OF THE WORK
  The open-source code \texttt{fdesign} makes it possible to design digital
  linear filters for the Hankel and Fourier transforms used in potential-,
  diffusive-, and wavefield modeling. Digital filters can be derived for any
  electromagnetic method, such as methods in the diffusive limits (DC, CSEM) as
  well as methods using higher frequency content (GPR, acoustic and elastic
  wavefields).
%
% (2) METHODOLOGY
  The direct matrix inversion method is used for the derivation of the filter
  values, and a brute-force inversion is carried out over the defined spacing
  and shifting values of the filter basis. Included or user-provided
  theoretical transform pairs are used for the inversion. Alternatively, one
  can provide 1D subsurface models that will be computed with a precise
  quadrature method using the EM modeler \texttt{empymod} to generate
  numerical transform pairs.
%
% (3) RESULTS
  The comparison of the presented 201\,pt filter with previously presented
  filters shows that it performs better for some standard CSEM cases. The
  derivation of a longer 2001\,pt filter for a GPR example with 250\,MHz centre
  frequency proves that the filter method works also for wave phenomena, not
  only for diffusive EM fields.
%
% (4) CONCLUSIONS
  The presented algorithm provides a tool to create problem specific digital
  filters. Such purpose-built filters can be made shorter and can speed up
  consecutive  potential-, diffusive-, and wavefield inversions.
%
\end{abstract}

\section{Introduction}

In his Ph.D. thesis, \cite{PhD.70.Ghosh} proposed a linear filter method for
the numerical evaluation of Hankel transforms that subsequently revolutionized
the computation of electromagnetic (EM) responses in the field of geophysical
exploration. If you use a code that calculates EM responses in the
wavenumber-frequency domain and transforms them to the space-frequency domain,
chances are high that it uses the \emph{digital linear filter} (DLF) method.
Most practical 1D EM modeling codes rely on the DLF method for rapid
computations; these codes are not only used for standalone simulations of EM
fields in layered 1D media, but they are also commonly embedded within 3D
modeling codes for computing the primary fields in scattered-field formulations
or for the background fields required by integral equation methods. Thus, the
DLF method is an important component of many commonly used modeling codes for
EM geophysical data.

\cite{GP.71.Ghosh} states that the DLF idea is based on suggestions made four
decades earlier by \cite{PHY.33.Slichter} and \cite{GEO.40.Pekeris}, in that
``\emph{the kernel function is dependent only on the layer parameters, and an
expression relating it to the field measurements can be obtained by
mathematical processes.}'' However, until the introduction of DLF, these
suggestions found no widespread use, likely because of the missing computer
power to calculate the filter coefficients. He further states that credit goes
to his Ph.D. supervisor Otto \cite{BK.68.Koefoed, GP.70.Koefoed}, who retook
the task of direct interpretational methods with the introduction of raised
kernel functions. DLF is, as such, an improvement of that approach, providing a
faster and simpler method.

The DLF technique is sometimes referred to as the \emph{fast Hankel transform}
(FHT), popular because of the similarity of the name to the well known
\emph{fast Fourier transform} (FFT), although the algorithms for these
techniques are completely different. The FHT name was likely inspired by the
title of a paper by \cite{GP.79.Johansen}. However, the name FHT can be
misleading as it has \emph{Hankel} in the name, whereas the DLF approach can
more generally be applied to other linear transforms, for example Fourier sine
and cosine transforms.

The introduction of linear filters to EM geophysics, in parallel with rising
computational power, initiated a wealth of investigations, leading to the
development of computer programs that extended and improved the DLF method, and
to numerous publications. These publications fall broadly into one or several
of three categories: (1) new applications that extend the usage of DLF to other
EM measurement techniques; (2) filter improvements that provide either new
filters or improved methods for the determination of filter coefficients; and
(3) computational tools that compute EM responses using DLF techniques. Here,
we briefly review the most relevant publications, without claiming
completeness.

\subsubsection{(1) New applications}

Ghosh used the method originally for the computation of type curves for
Schlumberger and Wenner resistivity soundings: \cite{GP.71.Ghosh} derives a
resistivity model from given Schlumberger or Wenner sounding curves; and
\citet{GP.71.Ghoshb} provides filters for the inverse operation, deriving
resistivity sounding curves from a given resistivity model. The method was next
applied to electromagnetic soundings with horizontal and perpendicular coils
\citep{GP.72.Koefoed}, to vertical coplanar coil systems \citep{GP.73.Verma},
to dipoles and other two electrode systems \citep{GP.74.Das, GP.74.Dasb,
GP.80.Das, GEO.94.Sorensen}, and to vertical dikes, hence vertical instead of
horizontal layers \citep{GEO.75.Niwas}. The first filters were specific to a
particular resistivity sounding type and its transform; later publications used
the method to get one type curve from another type curve \citep{GP.77.Kumar,
GP.78.Kumar}, or generalized the method to be applicable to a wider set of
problems \citep{EXG.80.Davis, GXP.81.Das, GEO.84.Das, GP.84.ONeill}.
Eventually, it passed from pure layered modeling to primary-secondary field
formulations for 3D problems, where DLF is used to compute the spatial
Fourier-Hankel transforms in a horizontally layered background medium and to
compute transient responses from frequency domain computations
\citep{GJI.81.Das, GJI.82.Das, GEO.84.Anderson, GEO.86.Newman,
MGS.17.Kruglyakov}. Other publications delved into the theory of the method,
analyzing the oscillating behaviour of the filters and trying to estimate the
error of DLF \citep{GP.72.Koefoedb, GP.76.Koefoedb, GP.79.Johansen,
GP.90.Christensen}.

\subsubsection{(2) Filter improvements}

\cite{PhD.70.Ghosh} derived the filter coefficients in the spectral domain by
dividing the output spectrum by the input spectrum followed by an inverse
Fourier transform. Improvements to the determination of filter coefficients
were provided by \cite{EXG.75.ONeill, GEO.77.Nyman, GEO.82.Das}, or
specifically for the Fourier transform by \cite{GP.86.Nissen}. A direct
integration method was used by \cite{GP.76.Bichara} and \cite{GP.78.Bernabini}.
\cite{GP.79.Koefoed} proposed a Wiener-Hopf least-squares method, which was
further improved by many authors \citep{GP.82.Guptasarma, GEO.1982.Murakami,
GP.97.Guptasarma}. \cite{GP.07.Kong} proposed a direct matrix inversion method
to solve the convolution equation, which requires only the input and output
sample values. To evaluate the filter's effectiveness, he defines a good filter
as one that recovers small or weak diffusive EM fields. This method was also
used by \cite{GEO.09.Key, GEO.12.Key} to create a suite of filters accurate for
marine EM data. Most works have published filters for the Hankel transform with
$J_0$ and $J_1$ Bessel functions (or $J_{-1/2}$, $J_{1/2}$ if applied to the
Fourier sine/cosine transform), as all higher Bessel functions can be
rewritten, via recurrence relations, using only use these two.
\cite{GP.94.Mohsen} is one of the rare cases which provides $J_2$ filter
weights.

\subsubsection{(3) Computational tools}

The most well-known codes are likely Anderson's freely available ones.
\cite{USGS.73.Anderson} extends the method to transient responses, applying DLF
not only to the Hankel transform, but also to the Fourier transform. A
transient signal can therefore be obtained be applying twice a digital filter
to the wavenumber-frequency domain calculation. \cite{USGS.75.Anderson,
GEO.79.Anderson} presents improved filters for both Fourier and Hankel
transforms, introducing measures to significantly speed-up the calculation,
such as the lagged convolution or using the same abscissae for $J_0$ and $J_1$.
\cite{TMS.82.Anderson} and the included 801\,pt filter became the de facto
industry standard, to which subsequent filters were compared.
\cite{GEO.89.Anderson} presents a hybrid solution that uses either the DLF or a
much slower but highly accurate adaptive-quadrature approach, which can also be
used to measure the accuracy of the DLF. \cite{GEO.12.Key} presented codes for
comparing the DLF approach with a speed optimized quadrature method call
quadrature-with-extrapolation (QWE); despite its optimizations, QWE is still
not as fast as DLF, but it remains valuable as an independent tool for testing
the accuracy of particular filters for the DLF method. Other examples include
the codes by \cite{GP.75.Johansen}, an interactive system for interpretation of
resistivity soundings, and a tool to calculate filter coefficients by
\cite{GP.90.Christensen}. The latter is available upon request and was used,
for instance, in all the open-source modeling and inversion routines of CSIRO
in the Amira Project 223 \citep{ASEG.07.Raiche}.

All mentioned publications have in common that they were derived for direct
current methods (DC) or low frequency methods, such as time-domain shallow EM
methods (TEM), or controlled-source electromagnetics (CSEM), but not for high
frequency methods such as ground-pe\-ne\-tra\-ting radar (GPR). Generally it
was even thought that the filter method works only in the diffusive limit where
the quasistatic approximation is valid \citep[e.g.,][]{GEO.15.Hunziker}.
Nevertheless, \cite{GEO.17.Werthmuller} tested DLF for modeling 250 MHz
center-frequency GPR data using a 401 point filter that was designed for
diffusive EM, and found that it was orders of magnitude faster than quadrature
approaches. However, good quality results were only obtained for the first
arrivals at short time-offsets and the later arrivals had poor quality. Yet,
this promising initial test suggested it might be possible to specifically
design a special filter for GPR frequencies where the EM fields propagate as
waves rather than by diffusion.

While there are several versions of digital filters that are freely available,
there are no open-source filter designing tools that readily available; they
are either only described methods in articles or the codes are only available
upon request by the author. Here, we address this gap by presenting the code
\texttt{fdesign}, which is an open-source tool for designing general or
purpose-built filters using the direct matrix inversion method. The code is
completely open-source and can be run using freely available programming
environments, thus there are few barriers to obtaining and running the code.

In the following sections we give a brief overview of the methodology, the
theory, and the code, and then show examples of its usage, as well as
presenting new filters for both CSEM and GPR data. The current article falls
therefore into all three mentioned categories by the new application of DLF to
wave phenomena and by providing new filters with the accompanying code. The
algorithm and many more examples of its usage can be found on
\href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
The examples can be used as templates to design new filters.

\section{Methodology}

The algorithm \texttt{fdesign} is a filter designing tool using the direct
matrix inversion method as described by \cite{GP.07.Kong} and based on scripts
developed by \cite{GEO.12.Key}. The tool is an add-on to the EM modeler
\texttt{empymod} \citep{GEO.17.Werthmuller}, written in Python, and hosted on
GitHub, which should foster interaction and enable anyone to toy around,
improve, and extend it. It can be used to derive digital linear filters for the
Hankel and the Fourier transforms for potential, diffusive, and wavefield
modeling (hence from DC to GPR).

Theoretically, it can be used to derive linear filters for any linear
transform, as long as it is supplied with the input and outputs of a transform
pair. Previous works have relied on using theoretical transforms pairs commonly
found in tables of integral transforms. Here we follow a similar approach, but
we also propose a new method where filters could be derived from accurately
computed numerical transform pairs, such as those computed using quadrature
approaches. For example, the EM modeler \texttt{empymod} could be used to
compute transform pairs for a particular class of model and data parameters
(for example, a specific EM method and model of interest) using a slow but
highly accurate quadrature method. A good filter derived from the resulting
transform pairs could then be used for faster modeling of similar model and
data parameters.

The main features of \texttt{fdesign} and some differences between it and the
methods it is based upon are:
\begin{itemize}
  \item The algorithm computes many different filters for various spacing and
    shift values (brute-force), with an optional minimization routine for
    the best outcome of the brute-force step.
  \item \cite{GP.07.Kong} and \cite{GEO.12.Key} optimize a filter for the $J_0$
    filter, and then use the obtained shift and spacing value to calculate the
    $J_1$ values. The presented algorithm can optimize either $J_0$, $J_1$, or
    both in the optimization process. Five transform-pairs for each $J_0$ and
    $J_1$, and three transform pairs for each sine and cosine of often used
    functions are included in the routine, but any other transform pair can be
    provided as input.
  \item More complex or more specific numerical transforms  pairs  can be
    provided instead of theoretical transform pairs by using the modeler
    \texttt{empymod}.
  \item \cite{GP.07.Kong} defines a \emph{good} filter as one that recovers
    \emph{weak} diffusive EM fields. In  \texttt{fdesign}, you can define a
    relative error level that defines up to what error an obtained result is
    good or not. The optimization seeks to find smallest amplitude field that
    is accurately modeled (i.e., minimizing the amplitude), but it has
    additionally another mode, where it will maximize the abscissae $r$ of the
    right-hand-side of the transform pair instead of minimizing the amplitude.
    The two modes yield the same result in the general, simple case of fast
    decaying transform pairs. However, for some complex models and specifically
    for very high frequencies, maximizing $r$ will yield much better and more
    consistent results.
  \item The algorithm can solve under-, equal-, and over-determined linear
    systems.
  \item The real or the imaginary part can be used for the inversion of complex
    transform pairs.
\end{itemize}

The quality of a filter depends heavily on the model chosen for comparison. An
obtained filter might be very good for one model or data acquisition setup, but
not that good for another one, which is no different for this designing tool.
There is no way to estimate the error of a result obtained with a certain
filter if you apply it to any other model, unless you calculate this other
model with another method for comparison. All results presented here should
therefore be taken with a certain care. Further, when applying a new filter to
a significantly different model and/or data space than what it has previously
been tested on, we recommend first verifying its accuracy in that space by a
comparison with results computed with a quadrature method such as QWE, which is
also available in \texttt{empymod}.

\subsection{Theory}

Most of the articles mentioned in the review have detailed derivations of the
digital filter method. In this article we focus on the algorithm, and summarize
the theory only very briefly by following \cite{GEO.12.Key}. In
electromagnetics we often have to evaluate integrals of the form
%
\begin{equation}
  F(r) = \int^\infty_0 f(l)K(l r)\mr{d}l \ ,
  \label{eq:HankelInt}
\end{equation}
%
where $l$ and $r$ denote evaluation values, and $K$ is the kernel function. In
the specific case of the Hankel transform $l$ corresponds to wavenumber, $r$ to
offset, and $K$ to Bessel functions; in the case of the Fourier transform $l$
corresponds to frequency, $r$ to time, and $K$ to sine or cosine functions. In
both cases it is an infinite integral which numerical integration is very
time-consuming because of the slow decay of the kernel function and its
oscillatory behaviour.

By substituting $r = e^x$ and $l = e^{-y}$ we get
%
\begin{equation}
  e^x F(e^x) = \int^\infty_{-\infty} f(e^{-y})K(e^{x-y})e^{x-y}\mr{d}y\ .
  \label{eq:filtint}
\end{equation}
%
This can be re-written as a convolution integral and be approximated for an
$N$-point filter by
%
\begin{equation}
  F(r) \approx \sum^N_{n=1} \frac{f(b_n/r) h_n}{r}\ ,
  \label{eq:filtapprox}
\end{equation}
%
where $h$ is the digital linear filter, and the logarithmically spaced filter
abscissae is a function of the spacing $\Delta$ and the shift $\delta$,
%
\begin{equation}
  b_n = \exp\left\{\Delta(-N/2+n) + \delta\right\} \ .
  \label{eq:base}
\end{equation}
%
From equation~\ref{eq:filtapprox} it can be seen that the filter method
requires $N$ evaluations at each $r$. For example, to calculate the frequency
domain result for 100 offsets with a 201\,pt filter requires 20,100 evaluations
in the wavenumber domain. This is why the DLF often uses interpolation to
minimize the required evaluations, either for $F(r)$ in what is referred to as
\emph{lagged convolution DLF}, or for $f(l)$, which we here call \emph{splined
DLF}.

In the direct matrix inversion method for solving for the digital filter
coefficients, equation \ref{eq:filtapprox} is cast as a linear system as
follows. $\Delta$ and $\delta$ are preassigned scalar values and $r$ is a range
of $M$ preassigned values $r_m$. The filter base coefficients $b_n$ are
computed using equation \ref{eq:base} and an array of values for $l$ are
computed using $l_{mn} = b_n/r_m$. The linear system's left hand side (LHS)
matrix $\bf{A}$ has dimensions $M \times N$ with coefficients $A_{mn} =
f(l_{mn})/r_n$ and the right hand side (RHS) vector $\bf{v}$ has elements $v_n
= F(r_n)$. The resulting linear system
%
\begin{equation}
  \bf{A}\bf{h} = \bf{v}
  \label{eq:,linearSystem}
\end{equation}
%
can be solved by direct matrix inversion to obtain the vector of filter
coefficients $\bf{h}$. For a given filter length $N$, there are many subjective
choices that go into the practical implementation of this method, including the
choice of the transform pairs $F(r)$ and $f(l)$, as well as the value for $M$
and the particular range and spacing of the values $r_m$. Once values are
chosen for these variables, an optimal filter can be found by a grid search
over $\Delta$ and $\delta$ for values that produce a high quality filter. The
choice of metric for what constitutes a high quality filter is also subjective
and is further discussed below.

\subsection{Pseudo-code for \texttt{fdesign} }

The main input variables are the filter length ($N$), the spacing ($\Delta$)
and shift ($\delta$) ranges over which to loop, and the transform pairs for the
inversion ($f_\mr{I}$) and the check of quality ($f_\mr{C}$). If $f_\mr{C}$ is
not provided, then $f_\mr{I}$ is used for both. There are additional, optional
input parameters, for instance to adjust how the RHS abscissae for the
inversion $r_\mr{I} = p(b, N)$ are calculated, where $b$ is the filter base.
Additionally, ranges $r$ for evaluating the check function need to be
specified.

The basic steps are as follows:
\begin{enumerate}
  \item Evaluate RHS of check-function $f_\mr{C}$:\newline
        $d_\mr{R} = f_\mr{C}.\mr{rhs}(r)$
  \item Loop over each value $\Delta_i$, $\delta_j$ (brute force):
    \begin{enumerate}
      \item Calculate filter base:\newline
        $b_n = \exp\left\{\Delta_i(-N/2+n) + \delta_j\right\}$
      \item Get required RHS and LHS evaluation points:\newline
        $r_\mr{I} = p(b, N)$\newline
        $l_\mr{I} = b/r_\mr{I}$
      \item Invert for filter coefficients:\newline
        $J_{ij} = \mr{solve}(f_\mr{I}.\mr{lhs}(l_\mr{I}),
        f_\mr{I}.\mr{rhs}(r_\mr{I}))$
    \item Calculate numerically RHS of check-function $f_\mr{C}$ with current
      filter $J_{ij}$:\newline
      $d_\mr{F} = f_\mr{C}.\mr{lhs}(b/r)\cdot J_{ij}\ /\ r$
      \item Store minimum recovered amplitude or maximum $r$ where relative
        error is less than the provided error:\newline
        $\chi_{ij} = g\left[
        \mr{argmin}\left(|(d_\mr{F}-d_\mr{R})/d_\mr{R}| >
        \mr{error}\right) -1\right]\ $, where $g$ is either $d_\mr{R}$
        or $1/r$.
    \end{enumerate}
  \item Return filter coefficients which yield minimum amplitude or maximum $r$
    (a local minimization can be run to polish the brute-force result):\newline
    return $J\left[\mr{argmin}(\chi)\right] $\newline
\end{enumerate}

\subsection{Minimization criteria}
Figure~\ref{fig:InvCrit} shows the differences between the different
minimization approaches; in (a) for a conventional, fast decaying transform
pair; and in (b) for a more complex, high frequency layered model. The circles
show the minimum amplitude used in previous approaches. This criteria is not
ideal, as it is subject to some random fluctuations and also depends on the
choice of $r$. The squares show the minimum amplitude given a certain
acceptable error, and the diamonds show the maximum $r$ given a certain
acceptable error (the inversion is a minimization process, it therefore
minimizes $1/r$, not $r$). In simple cases the minimum amplitude given an
acceptable error and the maximum $r$ given an acceptable error will yield the
same result. However, in complex cases the maximum $r$ is more consistent and
therefore a much better criteria.
%
\plot*{InvCrit}{width=.9\textwidth}{(a) A regular RHS curve of a transform pair
  with a purely decaying function. Minimum amplitude or maximum $r$ yield the
  same result in this scenario. Using a relative error criteria is more stable
  than just the absolute minimum amplitude. (b) A 1D model for $f=200\,$MHz.
  Here, the maximum $r$ provides a better criteria for the inversion.}
%

While the $\chi$ array is used to identify the optimal values of $\Delta$ and
$\delta$, the value of $\mr{argmin}(\chi)$ will depend on the choice of the
check function $f_C$ and can vary from one transform pair to another; thus it
should not be used to compare the quality of filters created with different
transform pairs.

\section{Numerical examples}

The numerical examples are focused on the Hankel transform, although
\texttt{fdesign} can be used to design digital linear filters for both Hankel
and Fourier transforms. As there is no difference in the procedure of the two,
this should be sufficient to demonstrate the algorithm.

\subsection{Design}
Figure \ref{fig:201} shows the solution spaces of four consecutive inversion
runs, where each run is a focus on a subsection of the previous run, indicated
by the red square.
%
\plot*{201}{width=\textwidth}{%
  Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
  consecutive run zooms into a portion of the previous solution space,
  indicated by the red square. The more in detail we obtain the solution space,
  the more random appears to be the distribution.}%
%
The filter length in this case is $N=201$, and the theoretical transform pairs
are as given, e.g., in \cite{USGS.75.Anderson}:
%
\begin{align}
  \int^\infty_0\,l \exp\left(-al^2\right) J_0(rl)\,dl &=
  \frac{\exp\left(\frac{-r^2}{4a}\right)}{2a}\ , \\
  \int^\infty_0\,l^2 \exp\left(-al^2\right) J_1(rl)\,dl &=
  \frac{r}{4a^2} \exp\left(-\frac{r^2}{4a}\right)\ ,
  \label{eq:j01}
\end{align}
%
where $a$ was set to 5. In the algorithm you can provide one pair for the
inversion, and a different pair to get the minimum amplitude or the maximum
$r$. In this case the same transform pair was used for both the inversion and
the check of quality. The acceptable relative error was set to 1\,\%. The RHS
evaluation parameter for the inversion was set so  that an over-determined
system was evaluated with $M = 2N$ equations, where $r_\mr{I}$ was
logarithmically spaced from $\log_{10}(1/\max(b)) - 1$ to $\log_{10}(1/\min(b))
+ 1$, where $b$ are the filter abscissae as given in equation~\ref{eq:base}.
This is accomplished in \texttt{fdesign} by setting parameter $r_\mr{def} = (1,
1, 2)$.

From the low resolution overview runs (a) and (b) it looks like this is a
standard, straight-forward minimization problem. However, from the more
detailed results in (c) and (d) it becomes obvious that it is a minimization
problem that has to be solved stochastically, as there are solutions with 2
orders of magnitude difference apparently randomly next to each other.

Figure \ref{fig:201b} shows in (a) the filter values for $J_0$ and $J_1$ of the
best 201 point filter obtained this way, and in (b) its right-hand-side
solution. The black dots indicate negative values, from which it can seen that
adjacent values are often alternating between positive and negative
contributions.
%
\plot*{201b}{width=.9\textwidth}{(a) Filter values of the best obtained 201\,pt
  filter with the corresponding check of quality in (b). Black points indicate
  negative values, which shows that adjacent values have often opposite signs.}
%

Designing filters is to a large extent trial and error. All input variables
influence the outcome, and often  a good filter is found by shear luck of
choosing the right starting parameters. Each of the input parameters has an
effect to the outcome. It also depends heavily on the transform pairs
$f_\mr{I;C}$; functions that decay rapidly are generally better, as noted by
earlier authors \citep[e.g.][]{USGS.75.Anderson}. It also depends on the filter
length, and obviously heavily on the spacing and shift values. Another
important point is how you define the right-hand evaluation points of the
inversion ($r_\mr{def}$). Evaluating corresponding transform pairs separately
or jointly also leads to different filter coefficients ($J_0$, $J_1$, or $J_0$
\& $J_1$, or equally sine, cosine, or sine \& cosine); also if the real or the
imaginary part is used when complex transform pairs or \texttt{empymod} is
used; and if it is inverted for the minimum amplitude or for the maximum $r$.

\subsection{CSEM}
In this section we compare the 201\,pt filter derived in the previous section
to marine CSEM models used in \cite{GP.07.Kong}, \cite{GEO.12.Key}, as well as
for a land EM case.

Figure~\ref{fig:kongf5} compares the derived 201\,pt filter with the two
half-spaces case used by \cite{GP.07.Kong} in his figure 5. The model consists
of a $0.3125\,\Omega\,$m sea layer of infinite thickness above a subsurface
$1\,\Omega\,$m half-space. The signal of a $1\,$Hz frequency $x$-directed
electric source 50\,m above the interface is measured at an $x$-directed
electric receiver at the interface.
%
\plot*{kongf5}{width=.9\textwidth}{(a) Results of different filters and QWE for
  the model of figure 5 from \cite{GP.07.Kong} with the relative errors shown
  in (b), using the QWE result. The relative error is meaningless from about
  15.5\,km onwards, as QWE failed as well for these very low amplitudes.}
%
In (a) it can be seen that the new 201\,pt filter is able to recover smaller
amplitudes than the 241\,pt filter from \cite{GP.07.Kong}, the 201\,pt filter
from \cite{GEO.12.Key}, and the 801\,pt filter from \cite{TMS.82.Anderson}
(Wer201, Kong241, Key201, and And801, respectively). It behaves equally well as
the quadrature with extrapolation (QWE), for which we used a 51\,pt quadrature
with relative and absolute tolerance of 1e-12 and 1e-30, respectively. The
relative error is shown in (b), where the QWE result was taken as \emph{truth}.
For offsets greater than roughly 15.5\,km the relative error becomes
meaningless, as the QWE fails itself; this part is greyed out in the figure.

Figure \ref{fig:keyf5}(a) considers the canonical CSEM model as used in
\cite{GEO.12.Key} in his figure 5: a $0.303\,\Omega\,$m sea with depth of
2\,km, a background resistivity of $1\,\Omega\,$m, in which a $100\,\Omega\,$m
target layer of 100\,m thickness is embedded at 1000\,m below the seafloor.
The source is located at 1990\,m and receivers are positioned on the seafloor.
%
\plot*{keyf5}{width=.9\textwidth}{(a) The canonical marine CSEM model of
  \cite{GEO.12.Key} and (b) a land model with source and receiver at the
  surface. The new filter has generally relative errors which are several
  orders of magnitude lower than the other filters.}
%

Figure \ref{fig:keyf5}(b) is a land case, with a background resistivity of
$10\,\Omega\,$m, in which a $500\,\Omega\,$m target of 100\,m thickness is
embedded at 1000\,m below the surface. The source depth is 0.5\,m and the
receiver depth is 0.8\,m. In both cases the new filter \emph{Wer201} has
generally a relative error which is orders of magnitude lower than the other
filters. The relative error of the real part is given in black, whereas the
relative error of the imaginary part is given in grey. It is interesting to
note that the real and imaginary parts have very similar errors in
\emph{Wer201}, \emph{Key201}, and \emph{And801}, but that the imaginary part of
\emph{Kong241} seems to behave considerably better for most part than its real
counterpart. To compare the real and the imaginary parts is insofar interesting
as the digital filters are purely real valued.

It is very important to note again that other scenarios might yield very
different error plots. Although the new 201\,pt filter proves to be very
accurate for these three models, it might not be the best filter in other
cases.  We also comment that while it is tempting to design very short length
filters (much less than about $N=100$) since shorter filters can greatly reduce
EM simulation times, in our experience, shorter filters tend to be less general
purpose than longer filters and we highly recommend thoroughly testing their
accuracy for the desired range of data and model parameters.

\subsection{Ground-penetrating radar}

Figure~\ref{fig:GPR} shows the GPR example as calculated in
\cite{GEO.15.Hunziker} and \cite{GEO.17.Werthmuller}; the model parameters are
given in subplot (a). The filter used for this example for the Hankel transform
is a 2001\,pt filter, derived with the fullspace solution with $f=500\,$MHz for
the inversion and the check of quality. The whole space solution used
resistivity $200\,\Omega\,$m, relative permittivity $\varepsilon_\mr{r}=10$,
and permeability $\mu_\mr{r}=1$. The source and receiver vertical separation is
1\,m. For the Fourier transform a 4096\,pt FFT was used with regularly spaced
frequencies from 0.5\,MHz to 850\,MHz and then zero-padded up to 2048\,MHz. The
frequency result is multiplied with a Ricker wavelet with a center frequency of
250\,MHz, and a gain function ($1 + |t^3|, t$ in ns) is applied. Subplot (b)
shows the result when using the adaptive quadrature (QUAD) option of
\texttt{empymod}, which uses the QAGSE routine from the Fortran QUADPACK
library. Subplot (c) shows the result using DLF with the new 2001\,pt digital
filter. The calculation with DLF took under 9 minutes and is therefore roughly
80 times faster than the calculation with QUAD, which took roughly 11 hours and
27 minutes. Calculating the same model with QWE took 7 hours and 20 minutes.
However, QWE uses in roughly 1/3 of the calculation internally QUAD in this
example, see \cite{GEO.17.Werthmuller}. (Note that DLF was run in parallel
using 4 threads at once, taking effectively only 2 minutes and 10 seconds to
calculate.
% EMmod \citep{GEO.15.Hunziker} took for the same model about 18 hours 55
% minutes, using a 61\,pt Gauss-Kronrod integration routine.
The lagged convolution version of DLF and the splined version of QWE were used
in this comparison.)
%
\plot*{GPR}{width=\textwidth}{GPR example for the model given in (a) using
  \texttt{empymod} with (b) the adaptive quadrature (QUAD) or with (c) a 2001\,pt digital linear filter (b). The
  quadrature took about 11\,h 27\,min to calculate, whereas the filter took
  2\,min 10\,s in parallel on 4 nodes, hence under 9\,min in total.}
%

Figure~\ref{fig:ftGPR} shows in (a)-(c) the real part of the frequency domain
results and in (d)-(f) the time domain results for offsets of 0.2\,m, 2.0\,m,
and 3.0\,m.
%
\plot*{ftGPR}{width=\textwidth}{The real-part frequency-domain and the
  time-domain responses for offsets of (a/d) 0.2\,m, (b/e) 2.0\,m, and (c/f)
  3.0\,m.}
%

The 2001\,pt filter was derived with the fullspace solution of a medium which
is very similar to the layer in which source and receiver reside in our GPR
example. To show that this filter can also be applied to different layer
parameters we run a test where we swapped layers one and two of the model in
Figure \ref{fig:GPR} (a). The resulting real and imaginary parts for the
offsets 0.2\,m, 2.0\,m, and 3.0\,m are given in Figure~\ref{fig:ftGPR2nd}.
%
\plot*{ftGPR2nd}{width=\textwidth}{The real and imaginary parts of the
  frequency-domain response for the model given in Figure~\ref{fig:GPR} (a)
  with swapped first and second layers for offsets of (a) 0.2\,m, (b) 2.0\,m,
  and (c) 3.0\,m.}
%

These examples show clearly that the filter method can indeed be applied to
high frequency EM modeling and therefore wave phenomena. We are convinced that
with further tests and analysis much better filters could be achieved, and
various concepts could be checked. One approach is to derive a filter for each
frequency band, say one for 1\,MHz--10\,MHz, one for 10\,MHz--100\,MHz, and one
for 100\,MHz--1\,GHz. Another approach could be to derive distinct filters for
$J_0$ and $J_1$, with different spacing and shift values. The first idea would
roughly triple the calculation cost, the second idea double them; however, they
both would still be very fast compared to standard quadrature methods. One
could also apply the digital filter method to acoustic and elastic wavefields
as well.

\section{Conclusions}

The presented, free and open-source code \texttt{fdesign} can be used to design
digital linear filters for the Hankel and Fourier transforms (and more
generally for any linear transform) using either analytical transform pairs or
1D subsurface models together with the EM-modeler \texttt{empymod}. The code is
available from GitHub in the add-on repository of \texttt{empymod}, called
\texttt{empyscripts}, from version 0.2.0 onwards.

The presented 201\,pt filter achieves more precise results in the three
presented CSEM cases than other filters, and is included in \texttt{empymod}
from version 1.4.5 onwards. However, as with any digital filter, the quality
depends heavily on the model, and this new filter might or might not behave
that well for other models or data parameters.

The GPR example shows that the digital linear filter method can also be used
for wave phenomena, not only for the diffusive approximation limit of low
frequency EM modeling.

We see \texttt{fdesign} as being useful for at least 3 scenarios:
\begin{enumerate}
  \item Providing a fast method to design problem-specific filters. For bigger
    inversion projects (such as, for instance, massive stochastic CSEM
    inversions) a purpose designed, short filter might save much more time than
    it costs to design it. This could even be integrated into inversion codes,
    as an optional pre-inversion step.
  \item Extending the filter method to new areas, namely higher frequencies and
    acoustic and elastic wavefields.
  \item Raising  interest for digital linear filters in geophysics by making it
    very easy for anyone to play around, create their own filters, and get a
    better understanding of the relative simplicity of the method. We are sure
    there are many great filters that are yet to be discovered.
\end{enumerate}

\section{Acknowledgment}

D.W. would like to thank the electromagnetic research group of Aleksandr
Mousatov at the IMP for fruitful discussions, and Jürg Hunziker for valuable
feedback which improved this manuscript considerably.

% REFERENCES
\begin{thebibliography}{}
\itemsep0pt

\bibitem[Anderson, 1973]{USGS.73.Anderson}
Anderson, W.~L.,  1973, Fortran {IV} programs for the determination of the
  transient tangential electric field and vertical magnetic field about a
  vertical magnetic dipole for an m-layer stratified earth by numerical
  integration and digital linear filtering: USGS, {\bf PB221240}.

\bibitem[Anderson, 1975]{USGS.75.Anderson}
--------, 1975, Improved digital filters for evaluating {F}ourier and {H}ankel
  transform integrals: USGS, {\bf PB242800}.
\newblock
  (\href{https://pubs.er.usgs.gov/publication/70045426}{https://pubs.er.usgs.gov/publication/70045426}).

\bibitem[Anderson, 1979]{GEO.79.Anderson}
--------, 1979, Numerical integration of related {H}ankel transforms of orders
  0 and 1 by adaptive digital filtering: Geophysics, {\bf 44}, 1287--1305.
\newblock (\href{http://doi.org/10.1190/1.1441007}{doi: 10.1190/1.1441007}).

\bibitem[Anderson, 1982]{TMS.82.Anderson}
--------, 1982, Fast {H}ankel transforms using related and lagged convolutions:
  ACM Trans. Math. Softw., {\bf 8}, 344--368.
\newblock (\href{http://doi.acm.org/10.1145/356012.356014}{doi:
  10.1145/356012.356014}).

\bibitem[Anderson, 1984]{GEO.84.Anderson}
--------, 1984, {On: “Numerical integration of related Hankel transforms by
  quadrature and continued fraction expansion” by \cite{GEO.83.Chave}}:
  Geophysics, {\bf 49}, 1811--1812.
\newblock (\href{http://doi.org/10.1190/1.1441595}{doi: 10.1190/1.1441595}).

\bibitem[Anderson, 1989]{GEO.89.Anderson}
--------, 1989, A hybrid fast {H}ankel transform algorithm for electromagnetic
  modeling: Geophysics, {\bf 54}, 263--266.
\newblock (\href{http://doi.org/10.1190/1.1442650}{doi: 10.1190/1.1442650}).

\bibitem[Bernabini and Cardarelli, 1978]{GP.78.Bernabini}
Bernabini, M., and E. Cardarelli,  1978, The use of filtered {B}essel functions
  in direct interpretation of geoelectrical soundings: Geophysical Prospecting,
  {\bf 26}, 841--852.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1978.tb01636.x}{doi:
  10.1111/j.1365-2478.1978.tb01636.x}).

\bibitem[Bichara and Lakshmanan, 1976]{GP.76.Bichara}
Bichara, M., and J. Lakshmanan,  1976, Fast automatic processing of resistivity
  soundings*: Geophysical Prospecting, {\bf 24}, 354--370.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1976.tb00932.x}{doi:
  10.1111/j.1365-2478.1976.tb00932.x}).

\bibitem[Chave, 1983]{GEO.83.Chave}
Chave, A.~D.,  1983, Numerical integration of related {H}ankel transforms by
  quadrature and continued fraction expansion: Geophysics, {\bf 48},
  1671--1686.
\newblock (\href{http://doi.org/10.1190/1.1441448}{doi: 10.1190/1.1441448}).

\bibitem[Christensen, 1990]{GP.90.Christensen}
Christensen, N.~B.,  1990, Optimized fast {H}ankel transform filters:
  Geophysical Prospecting, {\bf 38}, 545--568.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1990.tb01861.x}{doi:
  10.1111/j.1365-2478.1990.tb01861.x}).

\bibitem[Das, 1982]{GEO.82.Das}
Das, U.~C.,  1982, Designing digital linear filters for computing resistivity
  and electromagnetic sounding curves: Geophysics, {\bf 47}, 1456--1459.
\newblock (\href{https://doi.org/10.1190/1.1441295}{doi: 10.1190/1.1441295}).

\bibitem[Das, 1984]{GEO.84.Das}
--------, 1984, A single digital linear filter for computations in electrical
  methods--a unifying approach: Geophysics, {\bf 49}, 1115--1118.
\newblock (\href{https://doi.org/10.1190/1.1441726}{doi: 10.1190/1.1441726}).

\bibitem[Das and Ghosh, 1974]{GP.74.Das}
Das, U.~C., and D.~P. Ghosh,  1974, The determination of filter coefficients
  for the computation of standard curves for dipole resistivity sounding over
  layered earth by linear digital filtering: Geophysical Prospecting, {\bf 22},
  765--780.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1974.tb00117.x}{doi:
  10.1111/j.1365-2478.1974.tb00117.x}).

\bibitem[Das et~al., 1974]{GP.74.Dasb}
Das, U.~C., D.~P. Ghosh, and D.~T. Biewinga,  1974, Transformation of dipole
  resistivity sounding measurements over layered earth by linear digital
  filtering: Geophysical Prospecting, {\bf 22}, 476--489.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1974.tb00100.x}{doi:
  10.1111/j.1365-2478.1974.tb00100.x}).

\bibitem[Das and Verma, 1980]{GP.80.Das}
Das, U.~C., and S.~K. Verma,  1980, Digital linear filter for computing type
  curves for the two-electrode system of resistivity sounding: Geophysical
  Prospecting, {\bf 28}, 610--619.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1980.tb01246.x}{doi:
  10.1111/j.1365-2478.1980.tb01246.x}).

\bibitem[Das and Verma, 1981a]{GJI.81.Das}
--------, 1981a, Numerical considerations on computing the {EM} response of
  three-dimensional inhomogeneities in a layered earth: Geophysical Journal
  International, {\bf 66}, 733--740.
\newblock (\href{http://doi.org/10.1111/j.1365-246X.1981.tb04897.x}{doi:
  10.1111/j.1365-246X.1981.tb04897.x}).

\bibitem[Das and Verma, 1981b]{GXP.81.Das}
--------, 1981b, The versatility of digital linear filters used in computing
  resistivity and em sounding curves: Geoexploration, {\bf 18}, 297--310.
\newblock (\href{http://doi.org/10.1016/0016-7142(81)90059-4}{doi:
  10.1016/0016-7142(81)90059-4}).

\bibitem[Das and Verma, 1982]{GJI.82.Das}
--------, 1982, Electromagnetic response of an arbitrarily shaped
  three-dimensional conductor in a layered earth — numerical results:
  Geophysical Journal International, {\bf 69}, 55--66.
\newblock (\href{http://doi.org/10.1111/j.1365-246X.1982.tb04935.x}{doi:
  10.1111/j.1365-246X.1982.tb04935.x}).

\bibitem[Davis et~al., 1980]{EXG.80.Davis}
Davis, P.~A., S.~A. Greenhalgh, and N.~P. Merrick,  1980, Resistivity sounding
  computations with any array using a single digital filter: Exploration
  Geophysics, {\bf 11}, 54--62.
\newblock (\href{http://library.seg.org/doi/abs/10.1071/EG980054}{doi:
  10.1071/EG980054}).

\bibitem[Ghosh, 1970]{PhD.70.Ghosh}
Ghosh, D.~P.,  1970, The application of linear filter theory to the direct
  interpretation of geoelectrical resistivity measurements: {Ph.D. Thesis}, {TU
  Delft}.
\newblock
  (\href{http://resolver.tudelft.nl/uuid:88a568bb-ebee-4d7b-92df-6639b42da2b2}{uuid:
  88a568bb-ebee-4d7b-92df-6639b42da2b2}).

\bibitem[Ghosh, 1971a]{GP.71.Ghosh}
--------, 1971a, The application of linear filter theory to the direct
  interpretation of geoelectrical resistivity sounding measurements:
  Geophysical Prospecting, {\bf 19}, 192--217.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1971.tb00593.x}{doi:
  10.1111/j.1365-2478.1971.tb00593.x}).

\bibitem[Ghosh, 1971b]{GP.71.Ghoshb}
--------, 1971b, Inverse filter coefficients for the computation of apparent
  resistivity standard curves for a horizontally stratified earth: Geophysical
  Prospecting, {\bf 19}, 769--775.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1971.tb00915.x}{doi:
  10.1111/j.1365-2478.1971.tb00915.x}).

\bibitem[Guptasarma, 1982]{GP.82.Guptasarma}
Guptasarma, D.,  1982, Optimization of short digital linear filters for
  increased accuracy: Geophysical Prospecting, {\bf 30}, 501--514.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1982.tb01320.x}{doi:
  10.1111/j.1365-2478.1982.tb01320.x}).

\bibitem[Guptasarma and Singh, 1997]{GP.97.Guptasarma}
Guptasarma, D., and B. Singh,  1997, New digital linear filters for {H}ankel
  {J}0 and {J}1 transforms: Geophysical Prospecting, {\bf 45}, 745--762.
\newblock (\href{http://doi.org/10.1046/j.1365-2478.1997.500292.x}{doi:
  10.1046/10.1046/j.1365-2478.1997.500292.x}).

\bibitem[Hunziker et~al., 2015]{GEO.15.Hunziker}
Hunziker, J., J. Thorbecke, and E. Slob,  2015, The electromagnetic response in
  a layered vertical transverse isotropic medium: {A} new look at an old
  problem: Geophysics, {\bf 80}, no. 1, F1--F18.
\newblock (\href{http://doi.org/10.1190/geo2013-0411.1}{doi:
  10.1190/geo2013-0411.1}).

\bibitem[Johansen, 1975]{GP.75.Johansen}
Johansen, H.~K.,  1975, An interactive computer/graphic-display-terminal system
  for interpretation of resistivity soundings: Geophysical Prospecting, {\bf
  23}, 449--458.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1975.tb01541.x}{doi:
  10.1111/j.1365-2478.1975.tb01541.x}).

\bibitem[Johansen and Sørensen, 1979]{GP.79.Johansen}
Johansen, H.~K., and K. Sørensen,  1979, Fast {H}ankel transforms: Geophysical
  Prospecting, {\bf 27}, 876--901.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1979.tb01005.x}{doi:
  10.1111/j.1365-2478.1979.tb01005.x}).

\bibitem[Key, 2009]{GEO.09.Key}
Key, K.,  2009, {1D} inversion of multicomponent, multifrequency marine {CSEM}
  data: {M}ethodology and synthetic studies for resolving thin resistive
  layers: Geophysics, {\bf 74}, no. 2, F9--F20.
\newblock (\href{http://doi.org/10.1190/1.3058434}{doi: 10.1190/1.3058434}).

\bibitem[Key, 2012]{GEO.12.Key}
--------, 2012, Is the fast {H}ankel transform faster than quadrature?:
  Geophysics, {\bf 77}, no. 3, F21--F30.
\newblock (\href{http://doi.org/10.1190/GEO2011-0237.1}{doi:
  10.1190/GEO2011-0237.1}).

\bibitem[Koefoed, 1968]{BK.68.Koefoed}
Koefoed, O.,  1968, The application of the kernel function in interpreting
  geoelectrical resistivity measurements: Gebrüder Borntraeger.
\newblock ({I}SBN: 9783443130022).

\bibitem[Koefoed, 1970]{GP.70.Koefoed}
--------, 1970, A fast method for determining the layer distribution from the
  raised kernel function in geoelectrical sounding: Geophysical Prospecting,
  {\bf 18}, 564--570.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1970.tb02129.x}{doi:
  10.1111/j.1365-2478.1970.tb02129.x}).

\bibitem[Koefoed, 1972]{GP.72.Koefoedb}
--------, 1972, A note on the linear filter method of interpreting resistivity
  sounding data: Geophysical Prospecting, {\bf 20}, 403--405.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1972.tb00643.x}{doi:
  10.1111/j.1365-2478.1972.tb00643.x}).

\bibitem[Koefoed, 1976]{GP.76.Koefoedb}
--------, 1976, Error propagation and uncertainty in the interpretation of
  resistivity sounding data: Geophysical Prospecting, {\bf 24}, 31--48.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1976.tb00383.x}{doi:
  10.1111/j.1365-2478.1976.tb00383.x}).

\bibitem[Koefoed and Dirks, 1979]{GP.79.Koefoed}
Koefoed, O., and F.~J.~H. Dirks,  1979, Determination of resistivity sounding
  filters by the {W}iener-{H}opf least-squares method: Geophysical Prospecting,
  {\bf 27}, 245--250.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1979.tb00968.x}{doi:
  10.1111/j.1365-2478.1979.tb00968.x}).

\bibitem[Koefoed et~al., 1972]{GP.72.Koefoed}
Koefoed, O., D.~P. Ghosh, and G.~J. Polman,  1972, Computation of type curves
  for electromagnetic depth sounding with a horizontal transmitting coil by
  means of a digital linear filter: Geophysical Prospecting, {\bf 20},
  406--420.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1972.tb00644.x}{doi:
  10.1111/j.1365-2478.1972.tb00644.x}).

\bibitem[Kong, 2007]{GP.07.Kong}
Kong, F.~N.,  2007, Hankel transform filters for dipole antenna radiation in a
  conductive medium: Geophysical Prospecting, {\bf 55}, 83--89.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.2006.00585.x}{doi:
  10.1111/j.1365-2478.2006.00585.x}).

\bibitem[Kruglyakov and Bloshanskaya, 2017]{MGS.17.Kruglyakov}
Kruglyakov, M., and L. Bloshanskaya,  2017, High-performance parallel solver
  for integral equations of electromagnetics based on {G}alerkin method:
  Mathematical Geosciences, {\bf 49}, 751--776.
\newblock (\href{https://doi.org/10.1007/s11004-017-9677-y}{doi:
  10.1007/s11004-017-9677-y}).

\bibitem[Kumar and Das, 1977]{GP.77.Kumar}
Kumar, R., and U.~C. Das,  1977, Transformation of dipole to {S}chlumberger
  sounding curves by means of digital linear filters: Geophysical Prospecting,
  {\bf 25}, 780--789.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1977.tb01204.x}{doi:
  10.1111/j.1365-2478.1977.tb01204.x}).

\bibitem[Kumar and Das, 1978]{GP.78.Kumar}
--------, 1978, Transformation of {S}chlumberger apparent resistivity to dipole
  apparent resistivity over layered earth by the application of digital linear
  filters: Geophysical Prospecting, {\bf 26}, 352--358.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1978.tb01598.x}{doi:
  10.1111/j.1365-2478.1978.tb01598.x}).

\bibitem[Mohsen and Hashish, 1994]{GP.94.Mohsen}
Mohsen, A.~A., and E.~A. Hashish,  1994, The fast {H}ankel transform:
  Geophysical Prospecting, {\bf 42}, 131--139.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1994.tb00202.x}{doi:
  10.1111/j.1365-2478.1994.tb00202.x}).

\bibitem[Murakami and Uchida, 1982]{GEO.1982.Murakami}
Murakami, Y., and T. Uchida,  1982, Accuracy of the linear filter coefficients
  determined by the iteration of the least-squares method: Geophysics, {\bf
  47}, 244--256.
\newblock (\href{https://doi.org/10.1190/1.1441331}{doi: 10.1190/1.1441331}).

\bibitem[Newman et~al., 1986]{GEO.86.Newman}
Newman, G.~A., G.~W. Hohmann, and W.~L. Anderson,  1986, Transient
  electromagnetic response of a three-dimensional body in a layered earth:
  Geophysics, {\bf 51}, 1608--1627.
\newblock (\href{http://doi.org/10.1190/1.1442212}{doi: 10.1190/1.1442212}).

\bibitem[Nissen and Enmark, 1986]{GP.86.Nissen}
Nissen, J., and T. Enmark,  1986, An optimized digital filter for the {F}ourier
  transform: Geophysical Prospecting, {\bf 34}, 897--903.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1986.tb00500.x}{doi:
  10.1111/j.1365-2478.1986.tb00500.x}).

\bibitem[Niwas, 1975]{GEO.75.Niwas}
Niwas, S.,  1975, Direct interpretation of geoelectric measurements by the use
  of linear filter theory: Geophysics, {\bf 40}, 121--122.
\newblock (\href{https://doi.org/10.1190/1.1440510}{doi: 10.1190/1.1440510}).

\bibitem[Nyman and Landisman, 1977]{GEO.77.Nyman}
Nyman, D.~C., and M. Landisman,  1977, {VES} dipole-dipole filter coefficients:
  Geophysics, {\bf 42}, 1037--1044.
\newblock (\href{https://doi.org/10.1190/1.1440763}{doi: 10.1190/1.1440763}).

\bibitem[O'Neill, 1975]{EXG.75.ONeill}
O'Neill, D.,  1975, Improved linear filter coefficients for application in
  apparent resistivity computations: Exploration Geophysics, {\bf 6}, 104--109.
\newblock (\href{http://library.seg.org/doi/abs/10.1071/EG975104}{doi:
  10.1071/EG975104}).

\bibitem[O'Neill and Merrick, 1984]{GP.84.ONeill}
O'Neill, D.~J., and N.~P. Merrick,  1984, A digital linear filter for
  resistivity sounding with a generalized electrode array: Geophysical
  Prospecting, {\bf 32}, 105--123.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1984.tb00720.x}{doi:
  10.1111/j.1365-2478.1984.tb00720.x}).

\bibitem[Pekeris, 1940]{GEO.40.Pekeris}
Pekeris, C.~L.,  1940, Direct method of interpretation in resistivity
  prospecting: Geophysics, {\bf 5}, 31--42.
\newblock (\href{https://doi.org/10.1190/1.1441791}{doi: 10.1190/1.1441791}).

\bibitem[Raiche et~al., 2007]{ASEG.07.Raiche}
Raiche, A., F. Sugeng, and G. Wilson,  2007, Practical {3D} {EM} inversion -
  the {P223F} software suite: ASEG Technical Program Expanded Abstracts,  1--4.
\newblock (\href{http://doi.org/10.1071/ASEG2007ab114}{doi:
  10.1071/ASEG2007ab114}).

\bibitem[Slichter, 1933]{PHY.33.Slichter}
Slichter, L.~B.,  1933, The interpretation of the resistivity prospecting
  method for horizontal structures: Physics, {\bf 4}, 307--322.
\newblock (\href{http://doi.org/10.1063/1.1745198}{doi : 10.1063/1.1745198}).

\bibitem[Sørensen and Christensen, 1994]{GEO.94.Sorensen}
Sørensen, K.~I., and N.~B. Christensen,  1994, The fields from a finite
  electrical dipole—a new computational approach: Geophysics, {\bf 59},
  864--880.
\newblock (\href{https://doi.org/10.1190/1.1443646}{10.1190/1.1443646}).

\bibitem[Verma and Koefoed, 1973]{GP.73.Verma}
Verma, R.~K., and O. Koefoed,  1973, A note on the linear filter method of
  computing electromagnetic sounding curves: Geophysical Prospecting, {\bf 21},
  70--76.
\newblock (\href{http://doi.org/10.1111/j.1365-2478.1973.tb00015.x}{doi:
  10.1111/j.1365-2478.1973.tb00015.x}).

\bibitem[Werthmüller, 2017]{GEO.17.Werthmuller}
Werthmüller, D., 2017, An open-source full {3D} electromagnetic modeler for
  {1D} {VTI} media in {P}ython: empymod: Geophysics, {\bf 82}, no. 6,
  WB9--WB19.
\newblock (\href{http://doi.org/10.1190/geo2016-0626.1}{doi:
  10.1190/geo2016-0626.1}).

\end{thebibliography}

\end{document}